On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-08-25 Revisiones/Responses to the reviewers.tex

\documentclass[12pt]{article}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\geometry{a4paper} % or letter or a5paper or ... etc
\usepackage{graphicx}
\usepackage{cite}
\usepackage{color}
\usepackage{bm}

% \geometry{landscape} % rotated page geometry

\title{Response to the reviewers of the manuscript ``On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme"}
\author{Miguel Molinos, Pedro Navas, Manuel Pastor, and Miguel Mart\'in-Stickle}
\date{September 1, 2020}

%\date{} % delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}

\maketitle
%\tableofcontents

We  are grateful to the reviewers for taking their time  to review our work. Their comments have helped us to improve the paper. Changes to the original manuscript are given in  color (\textcolor{red}{red} for corrections, \textcolor{blue}{blue} for modified texts)  in the revised version.  A detailed response to both reviewers is given below.

\section*{Response to Reviewer \#1}
{\it
The authors have presented a promissing approach for MPM to surpass the shock wave propagation in the traditional MPM. The paper was very well written and the results suggested that the proposed approach works.
}
\\
\\
The reviewer's positive commentaries are greatly appreciated.
\\
\begin{enumerate}
\item \textit{However, most of the applications presented in this manuscript are relatively simple and does not really demonstrate the capability of the proposed approach. I suggest the authors to extend their manuscript to predict more challenging problems and compare their results to the original MPM solutions.}\\

Authors are completely agree with the proposal. And in analogy with the work of \cite{HAMMERQUIST2017724}, where fast dynamics were also studied. The Sulsky classical example of two colliding disks has been discussed in the new version of the manuscript.

\end{enumerate}

\hspace{5mm}

\section*{Response to Reviewer \#2}

\textit {This is, on the face of it, an interesting piece of work but in its present form I don't think this is ready for CMAME. The paper is too long, rambling and has many grammatical and spelling errors. }\\

Authors really appreciate that the reviewer consider the present research as an interest piece of work. In order to improve the readability of the revised paper, a proofreading is made.\\

\textit{Technical issues for the authors follow:}
 \begin{enumerate}
\item \textit{Cell-crossing is a well known issue with the MPM but in the review no mention is made of CPDI approaches.}\\

A citation of the work of \cite{Sadeghirad_2011} with the CPDI has been added to the state of art.

\item \textit{The development is very confusing and would make more sense were it to be tied to 
the way that GIMP methods are usually stated, i.e. that the standard FE shape functions are replaced by weighting functions comprised of the conjugation of a standard shape function and a particle "characteristic" function. Line 155 demonstrates this, i.e. the use of $N_{Ip}$ before explanation.}\\

The development of the MPM approach has been revised considering the proposal of the reviewer.

\item \textit{There is little discussion on some of the pitfalls of using max-ent, e.g. the need for a convex domain, the issue of determining Lagrange Multiplier solutions for points close to boundaries, i.e. the solutions process here at Eqns 37 and 38.}\\

Further discussion has been added.

\item \textit{Does the use of max-ent solve the problem of imposing Dirichlet boundary conditions in the MPM?}\\

No it does not. Local \textit{max-ent} approximation technique only provides to the MPM an efficient tool to transfer information from particles to nodes and vice versa. The difficulties of imposing Dirichlet boundary conditions (DBC) in the MPM are a direct consequence of discretising the continuum with material points nor with nodes where DBC are imposed. See for instance the work of \cite{Cortis_et_al_2017_IJNME}.

\item \textit{I am sceptical of the novelty of the proposed explicit scheme as there is no mention of the recent XPIC(m) approach of Hammerquist and Nairn or similar approaches used in the Computational Graphics community, e.g. Stomakhin et al. (2013) or the more recent Polypic.}\\

These publications will be added to the state of art of the manuscript. In the humble opinion of the authors, the XPIC approach consists in the stabilisation of the solution through a novel transference procedure. Meanwhile, the algorithm introduced in this research proposes the stabilisation within the adaptation of a FEM time integration scheme for the MPM framework. Hence, both techniques can be employed simultaneously.

\end{enumerate}

 \textit{The numerical examples are not impressive when the message of the paper is that this is good for fast dynamics problems. Compare to papers such as Ma et al. (2009) IJ Impact Engineering}\\

We agree with the reviewer that some of the numerical example does not enough highlights the capabilities of the MPM Newmark Predictor-Corrector scheme. Although the results presented in \cite{MA2009272} are impressive, the proposed examples are out of the scope of the present research. Therefore, in analogy with the work of \cite{HAMMERQUIST2017724}, previously cited by the reviewer, the Sulsky classical example of two colliding disks has been discussed.

\bibliographystyle{unsrt}
\begin{thebibliography}{10}

\bibitem{Cortis_et_al_2017_IJNME}
Cortis, Michael and Augarde, Charles and Coombs, William and Robinson, Scott and Brown, Michael and Brennan, Andrew.
\newblock {Imposition of essential boundary conditions in the material point method}.
\newblock {International Journal for Numerical Methods in Engineering}, 113, 2017.

\bibitem{Sadeghirad_2011}
Sadeghirad, A. and Brannon, R. M. and Burghardt, J.
\newblock {A convected particle domain interpolation technique to extend applicability of the material point method for problems involving massive deformations}.
\newblock {International Journal for Numerical Methods in Engineering}, 86(12) : 1435-1456, 2011.

\bibitem{HAMMERQUIST2017724}
Chad C. Hammerquist and John A. Nairn.
\newblock {A new method for material point method particle updates that reduces noise and enhances stability}.
\newblock {\em Computer Methods in Applied Mechanics and Engineering},  318:724--738, 2017.

\bibitem{MA2009272}
S. Ma and X. Zhang and X.M. Qiu.
\newblock{Comparison study of MPM and SPH in modeling hypervelocity impact problems}.
\newblock{International Journal of Impact Engineering}, 36(2):272--282, 2009.

\end{thebibliography}

\end{document}

%Del revisor 2:
%1- Meter paja de eso y fuera
%2- Yo haria referencia al paper de Wobbes, como lo nombren ellos, y decir que haces referencia a un paper contrastado....
%3- Hablar de eso, que hay pequeña discusion, pues se discute mas... 
%4- No, no lo soluciona, pero en OTM lo hace. Habría que conseguir una conjunción de lo que hace OTM con lo que hacen otros metodos en MPM para imponer boundary conditions....
%5- Leerse que hacen esos pollos, si esta bien y crees que puedes mter algun ejemplo, se puede intentar y asi tb se contenta al revisor 1

### 2020-08-26 LME-MPM.tex
Fuera comentarios
Hoja de respuestas revisores
Cambios en la derivación del modelo MPM

\usepackage{lineno}
\usepackage{hyperref}
\usepackage{float}
\usepackage{textcomp}
\usepackage{subfig}
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}

@@

\newcommand{\vect}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\tens}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Matrix}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Vector}[1]{
  \ensuremath{\mathbf{{#1}}}
}
% Divergence
\newcommand{\Div}[1]{
  \ensuremath{div({#1})}
}
% Gradient
\newcommand\Grad[1]{grad({#1})}
\newcommand\GradS[1]{grad^s({#1})}
\newcommand\GradT[1]{grad^T({#1})}
% Partial derivative
\newcommand{\Deriv}[3][]{
  \ensuremath{\frac{\partial^{#1}{#2}}{ \partial {#3}^{#1} }}
}
% Integral
\newcommand{\Integral}[2]{
  \IfStrEqCase{#1}{
    {2}{\ensuremath{\int_{\varGamma_d}{#2}\ d\varGamma}}
    {3}{\ensuremath{\int_{\varOmega}{#2}\ d\varOmega}}
  }
}

@@

\newglossaryentry{domain}
{
  name=$\varOmega$,
  description={Continuum domain}
}

\newglossaryentry{contour}
{
  name=$\partial\varOmega$,
  description={Boundary of the continuum domain $\varOmega$. Also
    defined as $\Gamma$}
}

\newglossaryentry{dirichlet-boundary}
{
  name=$ \Gamma_d$,
  description={Essential or Dirichlet boundary conditions over $\partial\varOmega$}
}

\newglossaryentry{neumann-boundary}
{
  name=$ \Gamma_n$,
  description={Natural or Neumann boundary conditions over $\partial\varOmega$}
}

@@

\newglossaryentry{a}
{
  name=$\vect{a}$,
  description={First order tensor which describes the acceleration field}
}

\newglossaryentry{v}
{
  name=$\vect{v}$,
  description={First order tensor which describes the velocity field}
}

\newglossaryentry{u}
{
  name=$\vect{u}$,
  description={First order tensor which describes the displacement field}
}

\newglossaryentry{x}
{
  name=$\vect{x}$,
  description={First order tensor which describes the global coordinates field}
}

\newglossaryentry{xi}
{
  name=$\vect{\xi}$,
  description={First order tensor which describes the local coordinates field}
}

@@

\newacronym{mpm}{MPM}{Material Point Method}
\newacronym{otm}{OTM}{Optimal Transportation Meshfree}
\newacronym{sph}{SPH}{Smoothed Particle Hydrodynamics}
\newacronym{fem}{FEM}{Finite Element Method}
\newacronym{efgm}{EFGM}{Element-Free Galerkin Method} 
\newacronym{lme}{LME}{Local Maximum-Entropy}
\newacronym{flip}{FLIP}{Fluid Implicit Particle}
\newacronym{gimp}{GIMP}{Generalized Interpolation Material Point}
\newacronym{igimp}{iGIMP}{Implicit GIMP}
\newacronym{ugimp}{uGIMP}{Uniform GIMP}
\newacronym{ddmp}{DDMP}{Dual Domain Material Point}
\newacronym{ctls}{CTLS}{Conservative Taylor Least Squares}
\newacronym{npc}{NPC}{Newmark Predictor-Corrector}
\newacronym{fe}{FE}{Forward Euler}

@@

\newcommand{\PNA}[1]{
  \textcolor{red}{{#1}}
}

\newcommand{\MMP}[1]{
  \textcolor{blue}{{#1}}
}

\newcommand{\MP}[1]{
  \textcolor{green}{{#1}}
}

@@

\journalname{Computer Methods in Applied Mechanics and Engineering}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

@@

%% Group authors per affiliation:
\author{
Miguel Molinos$^a$,
Pedro Navas$^a$\footnote{Corresponding author: p.navas@upm.es},
Manuel Pastor$^a$
and Miguel Mart\'in Stickle$^a$ 
}
\address{
  $^a$ ETSI Caminos, Canales y Puertos, Universidad Polit\'ectnica de Madrid.\\ c. Prof. Aranguren 3, 28040 Madrid, Spain
}

@@

\section*{-}
\centering
\Large
\textbf{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}\\
\setlength{\parskip}{1cm plus 5mm minus 4mm}
Miguel Molinos, Pedro Navas, Manuel Pastor and Miguel Mart\'in-Stickle.

@@

On the one hand, the main source of instability occurs when
material points cross cell boundaries. This provoked the development
of other interpolation techniques to overcome this limitation such as
the \acrfull{gimp} method Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrated to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is a voxel-based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed by Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the \acrfull{ddmp} method proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases. Therefore \acrshort{ddmp} a large
number of particles is needed for convergence~\cite{DHAKAL2016301}, what makes the method
very expensive. In recent years the employment of spline-lines as shape functions has gain popularity with the introduction of the B-Spline 
\acrshort{mpm} proposed by Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employment of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to the \acrshort{mpm} framework with the \acrfull{ctls}
reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}. Unfortunately, when particles are spread
in some particular special patterns, the quality of the \acrshort{ctls} approximation decreases locally.

@@

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers have proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing somehow the accuracy. One of the first attempts was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} mitigated these spurious noises by adding a non-viscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. In a recent publication,
Kularathna \& Soga \cite{Soga_2017} studied an implicit treatment of the pressure in MPM algorithm to simulate material incompressibility
avoiding artificial pressure oscillations by applying
Chorin\textquotesingle s
projection. Within the explicit time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort. In this paper a less time consuming and high efficient explicit predictor-corrector integration method is
proposed. It consists of an accommodation of the traditional \acrsull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

The aim of this document is to mitigate the spurious oscillations due to
inaccuracies in both space and time discretization by the employment of a
suitable combination of the \acrshort{lme} family shape functions, and the proposal of an explicit predictor-corrector scheme. The advantages of
this approach will be illustrated through several demanding test cases on the elastic regime: the
propagation of shock waves in an elastic bar and the response of a block of soil gradually loaded with gravitational forces.

@@

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method consists of three main steps: (i) a
variational recovery process, where particle data is projected onto the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention. Three kind of subscripts or superscripts are used
within paper. The subscript $\Box_p$ is used to define a particle
variable. While the subscript $\Box_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\Box^{\psi}$ involves a
virtual magnitude. For the operators, the convention is: $\dot{\Box}$ and
$\ddot{\Box}$ for the first and second time derivative, $\qued \otimes
\qued$ means the dyadic operator, $\Box \cdot \Box$ and $\Box \colon \Box$ means the
first and second contraction of a tensor, $\Div{\Box}$ denotes the
divergence operator, and finally $\Grad{\Box}$ and $\GradS{\Box}$
denotes the gradient and its symmetric part. Einstein subscripts
convention is adopted therefore repeated index means addition. Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are describe in subsection \ref{sec:derivation-mpm},\ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.

@@

In \acrshort{mpm} the continuum mechanics approach is considered. So on, we will consider a region \gls{domain} occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. First (i) we will relate
  global to local state by the \textit{compatibility equation}, where the the strain field \gls{strain} is extracted from \gls{u} is defined as:
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type
\gls{dirichlet-boundary}. We will further assume that strains are
infinitesimal, therefore second order terms
in the spatial derivatives can be neglected. From here, the
stress field \gls{stress} will be considered the corresponding
conjugate variable for the strain field, being the one which
satisfies (ii) the \textit{conservation of lineal momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type \gls{neumann-boundary}.  An aditional component will be (iii) the constitutive equation as a linear
application from $\mathbb{R}^n$ to $\mathbb{R}^n$, which relate stress and strain increments as,
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
In this research, plane strain Linear Elasticity has been
  considered. Thus the constitutive tensor, \tens{D}, is the well
  known linear elastic one. The final equation of the set is (iv) the mass
conservation, which can be obtained by setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

>>>>

\usepackage{lineno}
\usepackage{hyperref}
\usepackage{float}
\usepackage{textcomp}
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}

@@

\newcommand{\vect}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\tens}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Matrix}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Vector}[1]{
  \ensuremath{\mathbf{{#1}}}
}
% Divergence
\newcommand{\Div}[1]{
  \ensuremath{div({#1})}
}
% Gradient
\newcommand\Grad[1]{grad({#1})}
\newcommand\GradS[1]{grad^s({#1})}
\newcommand\GradT[1]{grad^T({#1})}
% Partial derivative
\newcommand{\Deriv}[3][]{
  \ensuremath{\frac{\partial^{#1}{#2}}{ \partial {#3}^{#1} }}
}
% Integral
\newcommand{\Integral}[2]{
  \IfStrEqCase{#1}{
    {2}{\ensuremath{\int_{\Gamma_d}{#2}\ d\Gamma}}
    {3}{\ensuremath{\int_{\Omega}{#2}\ d\Omega}}
  }
}

@@

\newglossaryentry{domain}
{
  name=$\Omega$,
  description={Continuum domain}
}

\newglossaryentry{contour}
{
  name=$\partial\Omega$,
  description={Boundary of the continuum domain $\Omega$. Also
    defined as $\Gamma$}
}

\newglossaryentry{dirichlet-boundary}
{
  name=$ \Gamma_d$,
  description={Essential or Dirichlet boundary conditions over $\partial\Omega$}
}

\newglossaryentry{neumann-boundary}
{
  name=$ \Gamma_n$,
  description={Natural or Neumann boundary conditions over $\partial\Omega$}
}

@@

\newglossaryentry{a}
{
  name=$\vec{a}$,
  description={First order tensor which describes the acceleration field}
}

\newglossaryentry{v}
{
  name=$\vec{v}$,
  description={First order tensor which describes the velocity field}
}

\newglossaryentry{u}
{
  name=$\vec{u}$,
  description={First order tensor which describes the displacement field}
}

\newglossaryentry{x}
{
  name=$\vec{x}$,
  description={First order tensor which describes the global coordinates field}
}

\newglossaryentry{xi}
{
  name=$\vec{\xi}$,
  description={First order tensor which describes the local coordinates field}
}

@@

\newacronym{mpm}{MPM}{Material Point Method}
\newacronym{otm}{OTM}{Optimal Transportation Meshfree}
\newacronym{sph}{SPH}{Smoothed Particle Hydrodynamics}
\newacronym{fem}{FEM}{Finite Element Method}
\newacronym{efgm}{EFGM}{Element-Free Galerkin Method} 
\newacronym{lme}{LME}{Local Maximum-Entropy}
\newacronym{flip}{FLIP}{Fluid Implicit Particle}
\newacronym{gimp}{GIMP}{Generalized Interpolation Material Point}
\newacronym{igimp}{iGIMP}{Implicit GIMP}
\newacronym{ugimp}{uGIMP}{Uniform GIMP}
\newacronym{ddmp}{DDMP}{Dual Domain Material Point}
\newacronym{cpdi}{CPDI}{Convected Particle Domain Interpolation}
\newacronym{ctls}{CTLS}{Conservative Taylor Least Squares}
\newacronym{npc}{NPC}{Newmark Predictor-Corrector}
\newacronym{fe}{FE}{Forward Euler}
\newacronym{dbc}{DBC}{Dirichlet boundary conditions}
\newacronym{nbc}{NBC}{Neumann boundary conditions}

@@

\newcommand{\CORRECTIONS}[1]{
  \textcolor{red}{{#1}}
}

\newcommand{\MODIFIED}[1]{
  \textcolor{blue}{{#1}}
}

@@

\journal{Computer Methods in Applied Mechanics and Engineering}
%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

@@

%% Group authors per affiliation:
\author{
Miguel Molinos$^a$,
Pedro Navas$^a$\footnote{Corresponding author: p.navas@upm.es},
Manuel Pastor$^a$
and Miguel Mart\'in Stickle$^a$}
\address{
  $^a$ ETSI Caminos, Canales y Puertos, Universidad Polit\'ectnica de Madrid.\\ c. Prof. Aranguren 3, 28040 Madrid, Spain
}

@@

\section*{}
\centering
\Large
\textbf{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}\\
\setlength{\parskip}{1cm plus 5mm minus 4mm}
Miguel Molinos, Pedro Navas, Manuel Pastor and Miguel Mart\'in-Stickle.

@@

On the one hand, the main source of instability occurs when
material points cross cell boundaries. This provoked the development
of other interpolation techniques to overcome this limitation such as
the \acrfull{gimp} method Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrated to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is a voxel-based
discretisation technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed by Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the \acrfull{ddmp} method proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases. Therefore \acrshort{ddmp} a large
number of particles is needed for convergence~\cite{DHAKAL2016301}, what makes the method very expensive. \MODIFIED{To avoid the tensile instabilities quite commons in extensions, Sadeghirad {\it et al.} (2011) \cite{Sadeghirad_et_al_IJNME_2011} developed the \acrshort{cpdi} \cite{Sadeghirad_et_al_IJNME_2011}}. In recent years the employment of spline-lines as shape functions has gain popularity with the introduction of the B-Spline 
\acrshort{mpm} proposed by Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employment of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to the \acrshort{mpm} framework with the \acrfull{ctls}
reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}. Unfortunately, when particles are spread
in some particular special patterns, the quality of the \acrshort{ctls} approximation decreases locally.

@@

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretisation technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers have proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing somehow the accuracy. One of the first attempts was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} mitigated these spurious noises by adding a non-viscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. In a recent publication,
Kularathna \& Soga \cite{Soga_2017} studied an implicit treatment of the pressure in MPM algorithm to simulate material incompressibility
avoiding artificial pressure oscillations by applying
Chorin\textquotesingle s
projection. Within the explicit time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort. In this paper a less time consuming and high efficient explicit predictor-corrector integration method is
proposed. It consists of an accommodation of the traditional \acrfull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

The aim of this document is to mitigate the spurious oscillations due to
inaccuracies in both space and time discretisation by the employment of a
suitable combination of the \acrshort{lme} family shape functions, and the proposal of an explicit predictor-corrector scheme. The advantages of
this approach will be illustrated through several demanding test cases on the elastic regime: the
propagation of shock waves in an elastic bar and the response of a block of soil gradually loaded with gravitational forces.

@@

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method consists of three main steps: (i) a
variational recovery process, where particle data is projected onto the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention. Three kind of subscripts or superscripts are used
within paper. The subscript $\square_p$ is used to define a particle
variable. While the subscript $\square_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\square^{\psi}$ involves a
virtual magnitude. For the operators, the convention is: $\dot{\square}$ and
$\ddot{\square}$ for the first and second time derivative, $\square \otimes
\square$ means the dyadic operator, $\square \cdot \square$ and $\square \colon \square$ means the
first and second contraction of a tensor, $\Div{\square}$ denotes the
divergence operator, and finally $\Grad{\square}$ and $\GradS{\square}$
denotes the gradient and its symmetric part. Einstein subscripts
convention is adopted therefore repeated index means addition. Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are describe in subsection \ref{sec:derivation-mpm},\ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.

@@

In \acrshort{mpm} the continuum mechanics approach is considered. So on, we will consider a region \gls{domain} occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \Omega
= \Gamma_d \cup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \Omega$
    of the domain $\Omega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. First (i) we will relate
  global to local state by the \textit{compatibility equation}, where the the strain field \gls{strain} is extracted from \gls{u} is defined as:
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vec{u}},
\end{equation}
together with essential or \acrfull{dbc} \gls{dirichlet-boundary}. We will further assume that 
strains are 
infinitesimal, therefore second order terms
in the spatial derivatives can be neglected. From here, the
stress field \gls{stress} will be considered the corresponding
conjugate variable for the strain field, being the one which
satisfies (ii) the \textit{conservation of lineal momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vec{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vec{b}
\end{equation}
together with the natural of \acrfull{nbc} \gls{neumann-boundary}.  An aditional component will be (iii) the constitutive equation as a linear
application from $\mathbb{R}^n$ to $\mathbb{R}^n$, which relate stress and strain increments as,
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
In this research, plane strain Linear Elasticity has been
  considered. Thus the constitutive tensor, \tens{D}, is the well
  known linear elastic one. The final equation of the set is (iv) the mass
conservation, which can be obtained by setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vec{v}} = 0.
\end{equation}
